### 2015-10-28 manuscript.Rnw
jwb edits

\documentclass[12pt,letterpaper]{article}

%%%%% packages provided by sys bio template
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{amsfonts}
\usepackage{verbatim}
%% \usepackage[english]{babel} %% I use polyglossia instead
%% not needed: \usepackage{pifont}
\usepackage{color}
\usepackage{setspace}
\usepackage{lscape}
\usepackage{indentfirst}
\usepackage[normalem]{ulem}
\usepackage{booktabs}
%\usepackage{nag}
\usepackage{natbib}
%\usepackage{bibtex}
\usepackage{float}
\usepackage{latexsym}
%\usepackage{hyperref}
\usepackage{url}
%\usepackage{html}
\usepackage{hyperref}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{array}
%\usepackage{mhchem}
\usepackage{ifthen}
%% I am already using it: \usepackage{caption}
\usepackage{hyperref}
%\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{amstext}
%%%%

% line numbers
\usepackage{lineno}
\modulolinenumbers[5]
\linenumbers

\usepackage{natbib}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage[font=sf]{caption}

% Latex special characters are rendered correctly with XeTeX
\usepackage{xltxtra}
\usepackage{xunicode}
\defaultfontfeatures{Mapping=tex-text}

% Words are cut where needed
\usepackage{polyglossia}
\setdefaultlanguage[variant=american]{english}

% Use fancy fonts
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{Crimson Text}
\setsansfont{SourceSansPro-Regular}

%%% syst bio options
\linespread{1.66}
% All text should be double-spaced
% with occasional exceptions for tables.
\raggedright
\setlength{\parindent}{0.5in}

\setcounter{secnumdepth}{0}

% Our sections are not numbered and our papers do not have
% Tables of Contents. We don't
% present a list of figures or list of tables, either.

% Any common font is fine.
% (A common sans-serif font should be used on figures, but figures should be
% separate from the LaTeX document.)

\pagestyle{empty}

\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}

\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}

\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}

\renewcommand{\tableofcontents}{}

\bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib
%%% end syst bio options

\newcolumntype{b}{X}
\newcolumntype{s}{>{\hsize=.2\hsize}X}

% ----------------------------------------------------------------------------- %

\begin{document}

\begin{flushright}
Version dated: \today
\end{flushright}

\bigskip
\noindent AN R PACKAGE TO INTERACT WITH THE OPEN TREE OF LIFE

\bigskip
\medskip

\begin{center}
  \noindent{\Large \bf \texttt{rotl} an R package to interact with the Open Tree
    of Life Data}

\bigskip

\noindent{\normalsize \sc Fran\c{c}ois Michonneau$^{1,2}$, Joseph Brown$^3$, David Winter$^4$}\\

\noindent {\small \it $^1$ Whitney Laboratory for Marine Sciences, University of
  Florida, St. Augustine, FL, USA}\\
\noindent {\small \it $^2$Florida Museum of Natural History, University of
  Florida, Gainesville, FL, USA} \\
\noindent{\small \it $^3$University of Michigan, Ann Arbor, MI, USA} \\
\noindent{\small \it $^4$Arizona State University, Tempe, AZ, USA} \\
\end{center}
\medskip

\noindent{\bf Corresponding author:} Fran\c{c}ois Michonneau, Division of
Invertebrate Zoology, Florida Museum of Natural History, Gainesville, FL
32611-7800, USA; E-mail: francois.michonneau@gmail.com\\

\vspace{1in}

\subsubsection{Abstract}

While phylogenies have been getting easier to build, it has been difficult to
re-use, combine, and synthesize the information they provide because published
trees are often only available as image files, and taxonomic information is not
standardized across studies. The Open Tree of Life (OTL) project addresses these
issues by providing a digital tree that encompasses all organisms built by
combining taxonomic information and published phylogenies. The project also
provides tools and services to query and download parts of this synthetic tree,
as well as the source data used to build it. Here, we present \texttt{rotl}, an R
package to search and download data from the Open Tree of Life directly in
R. \texttt{rotl} uses common data structures allowing researchers to take
advantage of the rich set of tools and methods that are available in R to
manipulate, analyze, and visualize phylogenies.

\vspace{1.5in}

Advances in sequencing and computing technologies have lead to a revolution in
systematic biology. The ability to routinely generate molecular datasets from
any extant organism has allowed researchers to resolve long-running taxonomic
disputes and estimate phylogenies for previously understudied groups. In
parallel, the ease with which phylogenies can be estimated has spurred the
development of novel phylogenetic comparative methods. These methods allow
researchers to explore fundamental questions about the origin of biodiversity,
including the evolution of morphological and ecological traits, the
spatio-temporal variation in speciation rates, or both
\citep{OMeara2012,Pennell2013}.

Ideally, the ever-increasing number of published phylogenies would contribute to
a synthesis of phylogenetic knowledge, ultimately leading to a richer understanding of
the history of life, while at the same time providing high-quality phylogenetic information for use
in comparative analyses. However, in practice synthesizing phylogenetic data
is a difficult task. Phylogenetic information is scattered, often only available as image
files within publications, and the lack of standardization to store and
represent phylogenetic data makes it difficult for researchers to access,
synthesize, and integrate this information into their own research
(\citealt{Stoltzfus2012}, \citealt{Drew2013}; but see \citealt{Cranston2014} for suggestions of best
practices).

The Open Tree of Life (OTL) project aims at assembling and synthesizing our
current understanding of phylogenetic relationships across all organisms on
Earth \citep{Hinchliff2015}, while providing tools and services that facilitate
access to this information. OTL combines taxonomic information that serves as
the backbone for the phylogenetic relationships, and published phylogenies to
elucidate relationships among taxa. This combination of information is used to
form a comprehensive synthetic tree. Studies can be contributed to the synthetic tree
through a curator interface (\url{https://tree.opentreeoflife.org/curator}),
allowing the synthetic tree to be continuously updated as relationships are
elucidated or reevaluated. The current draft of the OTL synthetic tree contains
2.3 million tips. Beyond obvious applications across the life sciences to
explore questions in evolution, biodiversity, and conservation, the resources
OTL provides is useful for education and outreach (e.g., illustrating course
material, or developing outreach activities to explore relationships among species).

% Because of its comprehensive nature, researchers have access to phylogenetic
% information for species that have yet to be included in any phylogenetic tree
% as their position is inferred based on taxonomy.

The OTL project provides a web interface to explore, query and
download parts of the synthetic tree (\url{http://tree.opentreeoflife.org})
as well as the published trees that contribute to it. Additionally, the project
provides Application Programming Interfaces (APIs)  to access programmatically
the different services and underlying data.  Here, we present \texttt{rotl}, a
package to interact with OTL's APIs to search for and import phylogenetic trees
and taxonomic information directly in R. Having access to these data in R allows
researchers to take advantage of the multiple packages that are available to
visualize, analyze, and manipulate phylogenetic and comparative data (e.g., ape
\citealt{Paradis2004}, phytools \citeateu{Revell2012}, geiger
\citealt{Pennell2014-geiger}, ggtree \citealt{Yu-ggtree}, phylobase
\citealt{phylobase-0.8.0}, RNeXML \citealt{Boettiger2015}; see
\url{https://cran.r-project.org/web/views/Phylogenetics.html} for a
comprehensive list). There is also a growing list of packages that allow users
to query and access data from the web directly in R: \hl{rFISHBASE
10.1111/j.1095-8649.2012.03464.x, paleoDB 10.1111/ecog.01154, rAvis
10.1371/journal.pone.0091650 \dots}. By providing direct access to high quality
phylogenetic data in R, \texttt{rotl} fills a key gap in typical comparative
analysis workflows. Moreover, by allowing the gathering, processing, and analysis
of data to be performed programmatically, \texttt{rotl} facilitates reproducible
research practices.

\section{API services provided by OTL}

The OTL project provides four resources that serve data to users through the
APIs:

\begin{enumerate}
\item The \emph{taxonomy} used as the backbone of the tree, the Open Tree
  Taxonomy (OTT);
\item The \emph{studies} and their associated trees, some of which are chosen by
  curators to assemble the synthetic tree;
\item A \emph{taxonomic name resolution service} (TNRS) used to match taxon
  names to the Open Tree Taxonomy identifiers;
\item The \emph{synthetic tree} itself, the `Open Tree'.
\end{enumerate}

Additionally, the project provides access to other data, the Graph of Life, that
stores information about the underlying tree alignment graphs \citep{Smith2013}
used to synthesize the various sources of information from which the Open Tree
is assembled. Data from this API are not intended for normal users, and its
use will not be demonstrated in this paper, although \texttt{rotl} has functions to
retrieve data it provides.

\texttt{rotl} gives users access to all endpoints provided by version 2 of the
OTL APIs. Phylogenetic trees served by the API can be imported directly into
R's memory and are represented using the \texttt{ape} \citep{Paradis2004} tree
structure (objects of class \texttt{phylo}), or can be written to files in the
Newick, NEXUS \citep{Maddison1997} or NeXML \citep{Vos2012} file formats. This
allows researchers to use these trees either directly with other R packages, or
to be imported in other programs that make use of phylogenetic tree files.

The full Open Tree currently does not have any branch lengths associated with
it, therefore parametric comparative methods cannot be used directly on the
subtrees returned by OTL (although the OTL treestore contains the raw 
published source phylogenies, complete with branch lengths and node 
annotations; see below). However, resources and methods are being developed to add
branch lengths to these topological subtrees \citep[e.g.,][]{Ksepka2015} or use
topological trees to identify phylogenetically equivalent species to increase
overlap between chronograms and species trait data \citep{Pennell2015}. Without
branch lengths, these subtrees are nonetheless useful to illustrate relationships
among species, or to map a trait on a phylogeny for instance.

%% any examples available?

% Tree versionning? Replicability of analyses

\section{Technical information about \texttt{rotl}}

Phylogenetic information retrieved from OTL is converted into
\texttt{ape::phylo} objects by \texttt{rotl} using the NEXUS Class Library
\citep[NCL,][]{Lewis2003} as implemented in the \texttt{rncl} package
(\url{https://cran.r-project.org/package=rncl}). Using NCL provides robust and
efficient parsing of large trees that may contain singleton nodes labeled with
taxonomic information (i.e., monotypic taxon).

The package is well-documented, and includes two package
vignettes (documents that demonstrate the use of the package and contain executable 
R code)
% as well as per-function documentation
. There is also an extensive
test-suite that covers both the internal functions that \texttt{rotl} uses to
connect to OTL and the public functions users apply to access and process data.

\section{Demonstrations}
\label{sec:demonstrations}

\subsection{Getting relationships from a list of taxa}
\label{sec:get-relationships}

To get the relationships among a set of taxa from the Open Tree, the taxa first
need to be matched against the OTL taxonomy OTT using the TNRS. This step retrieves
the identifiers that will be used to extract the relationships for the set of
requested taxa.

To illustrate how to obtain relationships from a set of taxa, we will rely on
OTL to draw a phylogenetic tree for a set of model organisms.
First, we use the function \texttt{tnrs\_match\_names} to match the taxon 
names to their Open Taxonomy identifiers.
<<init, echo=FALSE, purl=FALSE>>=
library(rotl)
@

%First, we use the function \texttt{tnrs\_match\_names} to match the taxon names to their Open Taxonomy identifiers.

<<get_taxa, cache=TRUE, purl=FALSE>>=
taxa <- tnrs_match_names(names = c("Escherichia colli",
                                   "Chlamydomonas reinhardtii",
                                   "Drosophila melanogaster",
                                   "Arabidopsis thaliana",
                                   "Rattus norvegicus",
                                   "Mus musculus",
                                   "Cavia porcellus",
                                   "Xenopus laevis",
                                   "Saccharomyces cervisae",
                                   "Danio rerio"))
@

The function \texttt{tnrs\_match\_names} returns a data frame that lists the
Open Tree identifiers as well as other information to help users ensure that the
taxa matched are the correct ones. Here, there is no ambiguity in the taxa
matched; however, as OTT includes taxa from bacteria, plants,
and animals that are regulated by different nomenclatural codes (ICNP, ICN, and
ICZN, respectively), both OTL and \texttt{rotl} provide tools to deal with potential
hemihomonyms. The argument \texttt{context\_name} can be used to limit potential
matches to a taxonomic group such as ``Animals'' (see the function
\texttt{tnrs\_contexts} for a complete list of possible options).  When this strategy
cannot be used (as in the present example, where the tree encompasses 
multiple domains), the function \texttt{inspect} lists alternative matches for a taxon
name, and \texttt{update} replaces it in the results. An example of this
approach is provided in the vignette ``How to use \texttt{rotl}?'' that
accompanies the package.

By default, approximate matching is enabled when attempting to match taxonomic
names to their OTT identifiers. Additionally, taxonomic synonyms
are included in OTT, allowing researchers to match correct
identifiers for taxon names that might include misspellings or synonyms. These
features will facilitate the tedious data cleaning process often needed when
mapping taxon names. In the example provided, both \textit{Escherichia coli}
and \textit{Saccharomyces cerevisiae} are misspelled, but the OTL TNRS finds the
correct match for these taxa.

Now that the taxon names are matched to the Open Tree identifiers, we can pass
them to the function \texttt{tol\_induced\_subtree} to retrieve the
relationships among these taxa. In turn, the tree can be plotted directly as it
is returned as an \texttt{ape::phylo} object (Figure~\ref{fig:plot_taxa}).

<<plot_taxa, fig.height=4, cache=TRUE, purl=FALSE, fig.cap="The phylogenetic tree returned by OTL for the list of model species used as an example.">>=
tree <- tol_induced_subtree(ott_ids = taxa[["ott_id"]])
plot(tree, cex=.8, label.offset=.1, no.margin=TRUE)
@

\subsection{Getting trees from studies}
\label{sec:get-tree-study}

\texttt{rotl} can also be used to retrieve trees accompanying studies that have
been submitted through the curator interface, and identify the trees that
contribute to the synthetic tree. These trees provide a useful resource to
reproduce or expand on a previously published analysis, or to explore how the
elucidation of relationships within a clade has changed through time.

Criteria that can be used to search for studies or their associated trees are
available through the output of the function
\texttt{studies\_properties}. Typically, users will want to search for studies
or trees based on taxon names (or their OTT identifiers), but other
criteria such as the title of the publication can be used. Here we demonstrate
how to look for and retrieve a tree for studies focusing on the family Felidae
(Figure~\ref{fig:plot_cats}).

<<get_cats, purl=FALSE, fig.cap="Tree from \\citealt{Johnson2006} obtained from OTL">>=
cat_studies <- studies_find_studies(property = "ot:focalCladeOTTTaxonName",
                                    value = "felidae")
cat_studies
@

Currently only one study focused on this family is available from OTL, and a
single tree is associated with it. We can then retrieve the study and tree
identifiers, and pass them to the function \texttt{get\_study\_tree} to have the
tree in memory:

<<get_cat_tree, purl=FALSE, results='hide'>>=
cat_tree <- get_study_tree(study_id = cat_studies[["study_ids"]][1],
                           tree_id = cat_studies[["tree_ids"]][1])
cat_tree
@

<<print_cat_tree, purl=FALSE, echo=FALSE>>=
print(cat_tree, printlen = 4)
@

<<plot_cats, purl=FALSE, fig.height=4.2, echo=FALSE, fig.cap='Phylogeny of the Felidae published in \\citealt{Johnson2006} and retrieved from OTL using \\texttt{rotl}'>>=
plot(cat_tree, label.offset=.0001, cex=.75, no.margin=TRUE)
@

When more than one tree is available for a given study, the function
\texttt{list\_trees} returns a list containing the tree identifiers for each
study. Alternatively, the function \texttt{get\_study} returns all the trees (by
default as \texttt{ape::phylo} objects) associated with a particular
study. Additional details about the study (i.e., its metadata) can be obtained
using the function \texttt{get\_study\_meta}.

\subsection{How does \texttt{rotl} fit into the R package ecosystem?}
\label{sec:how-does-rotl-fit}

In recent years, R has become an essential part of the toolbox of many researchers in
evolutionary biology. R greatly facilitates the analysis of large datasets, and
allows researchers to combine methods in novel ways because many methods for
comparative analyses are implemented, and because it is a relatively easy to use
programming language. Additionally, as more data are made available online and
accessible using APIs, several packages have been developed to interact and
download these datasets directly in R, enabling direct and reproducible analyses.
Notably, the organization rOpenSci (\url{https://ropensci.org}) has fostered a
community of researchers who develop tools and methods to facilitate the use of
open data, as well as broaden the adoption of open science practices in
general. \texttt{rotl} contributes to this initiative, and data from OTL can be
combined with other sources easily. For instance in Figure~\ref{fig:cat_map}, we
show how we can obtain a map of the occurrences for some of the cat species
(genus \textit{Lynx}) included in the phylogeny above using records for these
species found in GBIF (code available in the Appendix). Another example is available
in the package vignette ``Data mashups''.

<<cat_map, purl=TRUE, echo=FALSE, size='small', fig.height=4.2, results='hide', message=FALSE, cache=TRUE, fig.cap='GBIF records for the species in \\textit{Lynx} included in the phylogeny associated with the study by \\citealt{Johnson2006}'>>=
## load required packages
library(rotl)
library(rgbif)
library(ggplot2)
library(maps)

## Import the felid tree using study and tree IDs discovered with 
## studies_find_studies() in the manuscript

cat_tree <- get_study_tree(study_id ="pg_1981",                           
                           tree_id = "tree4052")

## Find the species of Lynx in the phylogeny
cat_species <- cat_tree$tip.label
lynx_species <- grep("^Lynx", cat_tree$tip.label, value = TRUE)

## Match the Lynx species to the GBIF identifiers
gbif_keys <- sapply(lynx_species,
                    function(x) name_backbone(name = x)$speciesKey,
                    USE.NAMES = FALSE)

## Search for the GBIF records for these species
lynx_loc <- occ_search(taxonKey = gbif_keys, limit = 500,
                       return = "data", fields = "minimal",
                       hasCoordinate = TRUE)

## Make a data frame of the results
lynx_loc <- do.call("rbind", lynx_loc)
names(lynx_loc)[1] <- "Species"

## Clean up the data with missing locality data
lynx_loc[["decimalLatitude"]] <- as.numeric(lynx_loc[["decimalLatitude"]])
lynx_loc[["decimalLongitude"]] <- as.numeric(lynx_loc[["decimalLongitude"]])
lynx_loc[lynx_loc[["decimalLatitude"]] == 0 &
         lynx_loc[["decimalLongitude"]] == 0,
         c("decimalLatitude","decimalLongitude")] <- c(NA, NA)
lynx_loc <- lynx_loc[complete.cases(lynx_loc), ]

## Draw the map
world <- map_data("world")

ggplot(lynx_loc) +
    annotation_map(world, fill="gray40", color="gray40") +
    geom_point(aes(y = decimalLatitude, x = decimalLongitude, color = Species),
               alpha = 0.3, size = 2) +
    coord_map(projection = "mercator", orientation = c(90, 0, 0)) +
    xlab("Longitude") + ylab("Latitude") +
    theme(legend.position="top") + ylim(c(0,72))

@

\section{Concluding remarks}
\label{sec:conclusion}

The recognition of the importance of phylogenies to account for the statistical
non-independence of species in comparative methods, and the recent development
of methods to explore trait evolution or changes in diversification rates, have
driven the need for accurate phylogenies. However, there is often a discrepancy
between taxa targeted by comparative methods, and taxa for which phylogenies are
available. We believe that by providing an easy-to-use interface to obtain
phylogenies for an arbitrary set of taxa directly in R, \texttt{rotl} will be
useful in a wide variety of contexts.

The accuracy and usefulness of the data provided by OTL relies on the community
to make generated phylogenies (and their metadata) digitally available as tree
files (i.e., Newick, NEXUS or NeXML). We strongly encourage researchers to
submit their published phylogenies to OTL using the curator interface
(\url{https://tree.opentreeoflife.org/curator}). By facilitating the discovery and
re-use of published trees and of the synthetic Open Tree, we hope \texttt{rotl} will
contribute to the wider adoption of best practices to make phylogenetic
information available and re-usable.

\section{Availability}
\label{sec:availability}

\texttt{rotl} is free, open source, and released under a Simplified
BSD license. Stable versions are available from the CRAN repository
(\url{https://cran.r-project.org/package=rotl}), and development
versions are available from GitHub
(\url{https://github.com/ropensci/rotl}). The package is under active development,
and authors welcome bug reports of feature requests via the github repository.
The source for this manuscript is available at \url{https://github.com/fmichonneau/rotl-ms}.

Python (\url{https://github.com/OpenTreeOfLife/pyopentree}) and Ruby
(\url{https://github.com/SpeciesFileGroup/bark}) libraries to interact
with the OTL APIs are also available.

\section{Acknowledgments}
\label{sec:acknowledgments}

We would like to thank the organizers of the OpenTree of Life APIs hackathon
that was held at the University of Michigan, Ann Arbor, 15-19 September, 2014,
where the development of \texttt{rotl} was started. We would also like to thank
Scott Chamberlain (rOpenSci) for providing a thorough code review. FM was
supported by iDigBio, and therefore this material is based upon work supported
by the National Science Foundation’s Advancing Digitization of Biodiversity
Collections Program (Cooperative Agreement EF-1115210)

\bibliographystyle{sysbio}

\bibliography{rotl-manuscript}

\end{document}