### 2016-03-30 manuscript.Rnw
reformat abstract according to guidelines

\documentclass[12pt,letterpaper]{article}

%%%%% packages provided by sys bio template
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{amsfonts}
\usepackage{verbatim}
%% \usepackage[english]{babel} %% I use polyglossia instead
%% not needed: \usepackage{pifont}
\usepackage{color}
\usepackage{setspace}
\usepackage{lscape}
\usepackage{indentfirst}
\usepackage[normalem]{ulem}
\usepackage{booktabs}
%\usepackage{nag}
\usepackage{natbib}
%\usepackage{bibtex}
\usepackage{float}
\usepackage{latexsym}
%\usepackage{hyperref}
\usepackage{url}
%\usepackage{html}
\usepackage{hyperref}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{array}
%\usepackage{mhchem}
\usepackage{ifthen}
%% I am already using it: \usepackage{caption}
\usepackage{hyperref}
%\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{amstext}
%%%%

%% Highlighting
\usepackage{soul}

% line numbers
\usepackage{lineno}
\modulolinenumbers[5]
\linenumbers

\usepackage{natbib}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage[font=sf]{caption}

% Latex special characters are rendered correctly with XeTeX
\usepackage{xltxtra}
\usepackage{xunicode}
\defaultfontfeatures{Mapping=tex-text}

% Words are cut where needed
\usepackage{polyglossia}
\setdefaultlanguage[variant=american]{english}

% Use fancy fonts
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{Crimson Text}
\setsansfont{SourceSansPro-Regular}

%%% syst bio options
\linespread{1.66}
% All text should be double-spaced
% with occasional exceptions for tables.
\raggedright
\setlength{\parindent}{0.5in}

\setcounter{secnumdepth}{0}

% Our sections are not numbered and our papers do not have
% Tables of Contents. We don't
% present a list of figures or list of tables, either.

% Any common font is fine.
% (A common sans-serif font should be used on figures, but figures should be
% separate from the LaTeX document.)

\pagestyle{empty}

\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}

\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}

\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}

\renewcommand{\tableofcontents}{}

\bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib
%%% end syst bio options

\newcolumntype{b}{X}
\newcolumntype{s}{>{\hsize=.2\hsize}X}

% ----------------------------------------------------------------------------- %

\begin{document}

\begin{flushright}
Version dated: \today
\end{flushright}

\bigskip
\noindent AN R PACKAGE TO INTERACT WITH THE OPEN TREE OF LIFE

\bigskip
\medskip

\begin{center}
  \noindent{\Large \bf \texttt{rotl} an R package to interact with the Open Tree
    of Life Data}

\bigskip

\noindent{\normalsize \sc Fran\c{c}ois Michonneau$^{1,2}$, Joseph Brown$^3$, David Winter$^4$}\\

\noindent {\small \it $^1$ Whitney Laboratory for Marine Sciences, University of
  Florida, St. Augustine, FL, USA}\\
\noindent {\small \it $^2$Florida Museum of Natural History, University of
  Florida, Gainesville, FL, USA} \\
\noindent{\small \it $^3$University of Michigan, Ann Arbor, MI, USA} \\
\noindent{\small \it $^4$Arizona State University, Tempe, AZ, USA} \\
\end{center}
\medskip

\noindent{\bf Corresponding author:} Fran\c{c}ois Michonneau, Division of
Invertebrate Zoology, Florida Museum of Natural History, Gainesville, FL
32611-7800, USA; E-mail: francois.michonneau@gmail.com\\

\vspace{1in}

\subsubsection{Abstract}

\begin{enumerate}
\item While phylogenies have been getting easier to build, it has been difficult
  to re-use, combine, and synthesize the information they provide because
  published trees are often only available as image files, and taxonomic
  information is not standardized across studies.
\item The Open Tree of Life (OTL) project addresses these issues by providing a
  digital tree that encompasses all organisms, built by combining taxonomic
  information and published phylogenies. The project also provides tools and
  services to query and download parts of this synthetic tree, as well as the
  source data used to build it. Here, we present \texttt{rotl}, an R package to
  search and download data from the Open Tree of Life directly in R.
\item \texttt{rotl} uses common data structures allowing researchers to take
  advantage of the rich set of tools and methods that are available in R to
  manipulate, analyze, and visualize phylogenies. Here, and in the vignettes
  accompanying the package, we demonstrate how \texttt{rotl} can be used with
  other R packages to analyze biodiversity data.
\item As phylogenies are being used in a growing number of applications,
  \texttt{rotl} facilitates access to phylogenetic data, and allows their
  integration with statistical methods and data sources available in R.
\end{enumerate}

\vspace{1.5in}

Advances in sequencing and computing technologies have lead to a revolution in
systematic biology. The ability to routinely generate molecular datasets from
any extant organism has allowed researchers to resolve long-standing taxonomic
disputes and estimate phylogenies for previously understudied groups. In
parallel, the ease with which phylogenies can be estimated has spurred the
development of new phylogenetic comparative methods. These methods allow
researchers to explore fundamental questions about the origin of biodiversity
including the evolution of morphological and ecological traits, the
spatio-temporal variation in speciation rates, or both
\citep{OMeara2012, Pennell2013}.

Ideally, the ever increasing number of published phylogenies would contribute to
a synthesis of phylogenetic knowledge, ultimately leading to a better
understanding of the history of life while at the same time providing
high-quality phylogenetic information for use in comparative analyses. However,
in practice, synthesizing phylogenetic data is a difficult task. Phylogenetic
information is largely scattered, often only available as image files within
publications, and the lack of standardization to store and represent
phylogenetic data makes it difficult for researchers to access, synthesize, and
integrate this information into their own research (\citealt{Stoltzfus2012,
  Drew2013, Magee2014}; but see \citealt{Cranston2014} for suggestions of best
practices).

The Open Tree of Life (OTL) project aims at assembling and synthesizing our
current understanding of phylogenetic relationships across all organisms on
Earth while providing tools and services that facilitate access to this
information \citep{Hinchliff2015}. OTL combines taxonomic information that
serves as the backbone for the phylogenetic relationships, and published
phylogenies to elucidate relationships among taxa. This combination of
information is used to structure the comprehensive synthetic tree. Studies can
be contributed to the synthetic tree through a curator interface
(\url{https://tree.opentreeoflife.org/curator}), allowing the synthetic tree to
be continuously updated as relationships are elucidated or reevaluated. The
current draft of the OTL synthetic tree contains 2.3 million tips. Beyond
obvious applications across the life sciences to explore questions in evolution,
biodiversity, and conservation, the resources OTL provides are useful for
education and outreach (e.g., illustrating course material, or developing
outreach activities to explore relationships among species).

% Because of its comprehensive nature, researchers have access to phylogenetic
% information for species that have yet to be included in any phylogenetic tree
% as their position is inferred based on taxonomy.

The R programming language is a popular tool for phylogenetics and comparative
analysis. The R packages \texttt{ape} \citep{Paradis2004}, \texttt{phylobase}
\citep{phylobase-0.8.0}, \texttt{phangorn} \citep{schliep2011} and
\texttt{RNeXML} \citep{Boettiger2015} each provide functions to import and
manipulate trees within R and save the results in standard data
formats. Additional packages including \texttt{phytools} \citep{Revell2012},
\texttt{geiger} \citep{Pennell2014-geiger}, and \texttt{ggtree}
\citep{Yu-ggtree}, allow users to analyze and visualize data in a phylogenetic
context (see \url{https://cran.r-project.org/web/views/Phylogenetics.html} for a
comprehensive list of phylogenetics packages in R). In addition to packages for
phylogenetic and comparative analyses, a growing number of R packages allows
users to query and access data from the web (e.g. \texttt{rFISHBASE}
\citep{Boettiger2012}, \texttt{rAvis} \citep{Varela2014}, and
\texttt{paleobioDB} \citep{Varela2015}), such that data associated with taxa in
a given phylogeny can be obtained directly in R.

In ecology, the development of the field of community phylogenetics
\citep{Webb2002} has created a need for researchers to have access to the
evolutionary relationships of species making up communities. The relative
contributions of the role of the environment (e.g., habitat filtering), and of
biotic interactions (e.g., competitive exclusion), are inferred from the
distribution of taxa on a phylogeny composed from species occurring at larger
spatial scale. R packages and other software have been developed to generate
phylogenies from species lists using taxonomic information or DNA sequences
\citep[e.g.,][]{Webb2008,Pearse2013}. These phylogenies can then be used for
community phylogenetics analyses \citep[e.g.,][]{Pearse2015, Kembel2010}, but
they are often incomplete or not resolved enough. As OTL becomes more
comprehensive, and its taxonomic resolution increases, it could become a
valuable resource for ecologists seeking to use phylogenetic information in
their research.

These packages, combined with the language's support for literate programming
(\citealt{Knuth1984}; \citealt{Xie2014}), make R a comprehensive platform for
reproducible research in phylogenetics and comparative biology, as they allow a
complete record of the steps taken in gathering, processing, and analyzing a
given data set to be produced.

Here we present \texttt{rotl}, an R package that allows users to download
phylogenetic and taxonomic data from the OTL directly in R. \texttt{rotl} takes
advantage of OTL's Application Programming Interfaces (APIs) to access sub-trees
from the synthetic Open Tree, as well as the published source trees that
contribute to the synthesis. By providing direct access to high quality
phylogenetic data in R, \texttt{rotl} fills a key gap in typical comparative
analysis workflows, and extends the degree to which R supports reproducible
research in phylogenetics and comparative biology.

\section{API services provided by OTL}

The OTL project provides four resources that serve data to users through the
APIs:

\begin{enumerate}
\item The \emph{taxonomy} used as the backbone of the tree, the Open Tree
  Taxonomy (OTT);
\item The \emph{studies} and their associated trees, some of which are chosen by
  curators to assemble the synthetic tree;
\item A \emph{taxonomic name resolution service} (TNRS) used to match taxon
  names to the Open Tree Taxonomy identifiers;
\item The \emph{synthetic tree} itself, the `Open Tree'.
\end{enumerate}

\texttt{rotl} gives users access to the endpoints provided by version 2 of the
APIs, and other versions of the APIs can be selected by the user as they become
available (version 3 is scheduled to be released at the end of March 2016). Two
other end-points are available in version 2 of the API: \emph{graph of life} and
\emph{collections}. Data from the \emph{graph of life} endpoint are not intended
for typical users, and will be removed in future releases. The \emph{collection}
endpoint is not fully functional, and no functions are available in the current
release of \texttt{rotl}. Future versions will make them available when the API
implements them.

Phylogenetic trees served by the API can be imported directly into R's
memory and are represented using the \texttt{ape} \citep{Paradis2004} tree
structure (objects of class \texttt{phylo}), or can be written to files in the
Newick, NEXUS \citep{Maddison1997}, or NeXML \citep{Vos2012} file formats. This
allows researchers to use these trees either directly with other R packages, or
to be imported in other programs that make use of phylogenetic tree files.

Currently, the synthetic tree does not have any branch lengths associated with
it, therefore parametric comparative methods cannot be used directly on the
subtrees returned by OTL (although the OTL treestore contains the raw published
source phylogenies, complete with branch lengths and node annotations; see
below). However, resources and methods are being developed to add branch lengths
to these topological subtrees \citep[e.g.,][]{Ksepka2015} or use topological
trees to identify phylogenetically equivalent species to increase overlap
between chronograms and species trait data \citep{Pennell2015}. Without branch
lengths, these subtrees are nonetheless useful to illustrate relationships among
species, or to map traits on a phylogeny.

%% any examples available?

% Tree versioning? Replicability of analyses

\section{Technical information about \texttt{rotl}}

Phylogenetic information retrieved from OTL is converted into \texttt{phylo}
objects by \texttt{rotl} using the NEXUS Class Library \citep[NCL,][]{Lewis2003}
as implemented in the \texttt{rncl} package
(\url{https://cran.r-project.org/package=rncl}). Using NCL provides robust and
efficient parsing of large trees that may contain singleton nodes labeled with
taxonomic information (i.e., a monotypic taxon). Singleton nodes are collapsed
after the tree has been parsed, making the resulting \texttt{phylo} object
compatible with all functions from the \texttt{ape} package.

The package is well-documented, and includes three package vignettes (documents
that demonstrate the use of the package and contain executable R code). There is
also an extensive test-suite that covers both the internal functions that
\texttt{rotl} uses to connect to OTL, and public functions that users apply to
access and process data.

\section{Demonstrations}
\label{sec:demonstrations}

\subsection{Getting relationships from a list of taxa}
\label{sec:get-relationships}

Before a researcher can use the Open Tree to retrieve relationships among a set
of taxa, they first need to match the taxon names in their dataset with records
in the Open Tree Taxonomy (OTT). OTL's taxonomic names resolution service (TNRS)
combines information from multiple services (e.g., National Center for
Biotechnology Information [NCBI], World Register for Marine Species [WoRMS],
Global Biodiversity Information Facility [GBIF]) and allows users to search for
taxon names, and retreive identifiers for each matching taxon. We demonstrate
the use of the TNRS within \texttt{rotl} by searching for taxonomic records
associated with several model organisms.

<<init, echo=FALSE, purl=FALSE>>=
library(rotl)
@
<<get_taxa, cache=TRUE, purl=FALSE>>=
taxa <- tnrs_match_names(names = c("Escherichia colli",
                                   "Chlamydomonas reinhardtii",
                                   "Drosophila melanogaster",
                                   "Arabidopsis thaliana",
                                   "Rattus norvegicus",
                                   "Mus musculus",
                                   "Cavia porcellus",
                                   "Xenopus laevis",
                                   "Saccharomyces cervisae",
                                   "Danio rerio"))
@

The function \texttt{tnrs\_match\_names} returns a data frame that lists the
Open Tree identifiers as well as other information to help users ensure that the
taxa matched are the correct ones. Here, there is no ambiguity in the taxa
matched; however, as OTT includes taxa from bacteria, plants, and animals that
are regulated by different nomenclatural codes (ICNP, ICN, and ICZN,
respectively), both OTL and \texttt{rotl} provide tools to deal with names that
may represent valid taxa in more than one code. The argument
\texttt{context\_name} can be used to limit potential matches to a taxonomic
group such as ``Animals'' (see the function \texttt{tnrs\_contexts} for a
complete list of possible options). When this strategy cannot be used (as in the
present example, where the tree encompasses multiple domains), the function
\texttt{inspect} lists alternative matches for a taxon name, and \texttt{update}
replaces it in the results. An example of this approach is provided in the
vignette ``How to use \texttt{rotl}?'' that accompanies the package.

By default, approximate matching is enabled when attempting to match taxonomic
names to their OTT identifiers. Additionally, taxonomic synonyms are included in
OTT, allowing researchers to match correct identifiers for taxon names that
might include misspellings or synonyms. These features will facilitate the
tedious data cleaning process often needed when mapping taxon names. In the
example provided, both \textit{Escherichia coli} and \textit{Saccharomyces
  cerevisiae} are misspelled, but the OTL TNRS finds the correct match for these
taxa.

Now that the taxon names are matched to the Open Tree identifiers, we can pass
them to the function \texttt{tol\_induced\_subtree} to retrieve the
relationships among these taxa. In turn, the tree can be plotted directly as it
is returned as \texttt{phylo} object (Figure~\ref{fig:plot_taxa}).

<<plot_taxa, fig.height=4, cache=TRUE, purl=FALSE, fig.cap="The phylogenetic tree returned by OTL for the list of model species used as an example.">>=
tree <- tol_induced_subtree(ott_ids = taxa[["ott_id"]])
plot(tree, cex = .8, label.offset = .1, no.margin = TRUE)
@

\subsection{Getting trees from studies}
\label{sec:get-tree-study}

\texttt{rotl} can also be used to retrieve trees accompanying studies that have
been submitted through the curator interface, and identify the trees that
contribute to the synthetic tree. As of March 2016, the Open Tree of Life
project stores 7755 trees from 3399 studies (each having between 0 and 61
trees), and 477 of these trees are used to assemble the synthetic tree. These
trees constitute a useful resource to reproduce or expand on a previously
published analysis, or to explore how the elucidation of relationships within
a clade has changed through time.

Criteria that can be used to search for studies or their associated trees are
available through the output of the function
\texttt{studies\_properties}. The meaning of these properties is described at:
\url{https://github.com/OpenTreeOfLife/phylesystem-api/wiki/NexSON}. Typically,
users will want to search for studies or trees based on taxon names (or their
OTT identifiers), but other criteria such as the title of the publication can be
used. Here we demonstrate how to look for and retrieve trees for studies
focusing on the family Felidae (Figure~\ref{fig:plot_cats}).

<<get_cats, purl=FALSE, fig.cap="Tree from \\citealt{Johnson2006} obtained from OTL">>=
cat_studies <- studies_find_studies(property = "ot:focalCladeOTTTaxonName",
                                    value = "Felidae", exact = TRUE)
cat_studies
@

Currently only one study focused on this family is available from OTL, and a
single tree is associated with it. We can then retrieve the study and tree
identifiers, and pass them to the function \texttt{get\_study\_tree} to have the
tree in memory:

<<get_cat_tree, purl=FALSE, results='hide'>>=
cat_tree <- get_study_tree(study_id = cat_studies[["study_ids"]][1],
                           tree_id = cat_studies[["tree_ids"]][1])
cat_tree
@

<<print_cat_tree, purl=FALSE, echo=FALSE>>=
print(cat_tree, printlen = 4)
@

<<plot_cats, purl=FALSE, fig.height=4.2, echo=FALSE, fig.cap='Phylogeny of the Felidae published in \\citealt{Johnson2006} and retrieved from OTL using \\texttt{rotl}.'>>=
plot(cat_tree, label.offset=.0001, cex=.75, no.margin=TRUE)
@

When more than one tree is available for a given study, the function
\texttt{list\_trees} returns a list containing the tree identifiers for each
study. Alternatively, the function \texttt{get\_study} returns all the trees (by
default as \texttt{phylo} objects) associated with a particular study. Metadata
about the study (e.g., citation information, information about the curator for
the study, and other technical information regarding the import of this study)
can be obtained using the function \texttt{get\_study\_meta}.

\subsection{How does \texttt{rotl} fit into the R package ecosystem?}
\label{sec:how-does-rotl-fit}

In recent years, R has become an essential part of the toolbox of many
researchers in evolutionary biology and ecology. R greatly facilitates the
analysis of large datasets, and allows researchers to combine methods in novel
ways because many methods for comparative analyses are implemented, and because
it is a relatively easy to use programming language. Additionally, as more data
are made available online and accessible using APIs, several packages have been
developed to interact and download these datasets directly in R, thereby
enabling direct and reproducible analyses.  Notably, the organization rOpenSci
(\url{https://ropensci.org}) has fostered a community of researchers who develop
tools and methods to facilitate the use of open data as well broaden the
adoption of open science practices in general \citep{Boettiger2015a}. For
instance, the rOpenSci-developed package \texttt{TreeBase}
\citep{Boettiger2012TB} allows users to access phylogenies stored in treeBASE
(\url{www.treebase.org}).  \texttt{rotl} contributes to this initiative, and
greatly extends the number of taxa for which phylogenetic data can be retrieved
within R, while allowing the data from OTL to be combined with other sources
easily.

Here we show how we can obtain a map of the occurrences for a subset of the cat
species that were included in the phylogeny retrieved from the Felidae study
above (genus \textit{Lynx}). We extract the species names from the phylogeny,
and use them to ask for the records for these species found in GBIF
(Figure~\ref{fig:cat_map}). We include the code to reproduce this figure in
Appendix.

As trait databases are becoming increasingly more common, and interfaces to the
data they contains are being developed as R packages \citep[e.g., the package
\texttt{traits}][]{Chamberlain2016}, \texttt{rotl} provides a way to easily
retrieve phylogenetic information for species trait data that are available.

In addition of an introduction on how to use the package, \texttt{rotl} includes
two vignettes that demonstrate how to integrate a phylogeny and data associated
with the taxa it represents. Specifically, the ``Data mashups'' vignette
provides an example of how to retrieve a phylogeny for species a researcher may
have data for, and visualize both the phylogeny and data associated with the
species at the tips. The other vignette titled ``Meta-analysis'' demonstrates
how a complete comparative method analysis, including the gathering of data and
a phylogeny, can be performed in a single R session. We reproduce a published
meta-analysis testing for differential investment in male and female offsprings
among 51 species of birds. As new versions of the OTL API and \texttt{rotl} are
released, these vignettes will be kept up-to-date.

<<cat_map, purl=TRUE, echo=FALSE, size='small', fig.height=4.2, results='hide', message=FALSE, cache=TRUE, fig.cap='GBIF records for the species in \\textit{Lynx} included in the phylogeny associated with the study by \\citealt{Johnson2006}.'>>=
## load required packages. Versions used for the manuscript listed below.
library(rotl)     # v0.5.0.901
library(rgbif)    # v0.9.2
library(ggplot2)  # v2.1.0
library(maps)     # v3.1.0

## Import the felid tree using study and tree IDs discovered with
## studies_find_studies() in the manuscript
cat_tree <- get_study_tree(study_id ="pg_1981",
                           tree_id = "tree4052")

## Find the species of Lynx in the phylogeny
cat_species <- cat_tree$tip.label
lynx_species <- grep("^Lynx", cat_tree$tip.label, value = TRUE)

## Match the Lynx species to the GBIF identifiers
gbif_keys <- sapply(lynx_species,
                    function(x) name_backbone(name = x)$speciesKey,
                    USE.NAMES = FALSE)

## Search for the GBIF records for these species
lynx_loc <- occ_search(taxonKey = gbif_keys, limit = 500,
                       return = "data", fields = "minimal",
                       hasCoordinate = TRUE)

## Make a data frame of the results
lynx_loc <- do.call("rbind", lynx_loc)
names(lynx_loc)[1] <- "Species"

## Clean up the data with missing locality data
lynx_loc[["decimalLatitude"]] <- as.numeric(lynx_loc[["decimalLatitude"]])
lynx_loc[["decimalLongitude"]] <- as.numeric(lynx_loc[["decimalLongitude"]])
lynx_loc[lynx_loc[["decimalLatitude"]] == 0 &
         lynx_loc[["decimalLongitude"]] == 0,
         c("decimalLatitude","decimalLongitude")] <- c(NA, NA)
lynx_loc <- lynx_loc[complete.cases(lynx_loc), ]

## Draw the map
world <- map_data("world")

ggplot(lynx_loc) +
    annotation_map(world, fill="gray40", color="gray40") +
    geom_point(aes(y = decimalLatitude, x = decimalLongitude, color = Species),
               size = 1) +
    coord_map(projection = "mercator", orientation = c(90, 0, 0)) +
    xlab("Longitude") + ylab("Latitude") +
    theme(legend.position="top", legend.key = element_rect(fill = "gray40")) +
    ylim(c(0,72))
@

\section{Concluding remarks}
\label{sec:conclusion}

The recognition of the importance of phylogenies to account for the statistical
non-independence of species in comparative methods, the recent development of
methods to explore trait evolution or changes in diversification rates, and
attempts to incorporate the evolutionary history of species forming ecological
communities, have driven the need for accurate phylogenies. However, there is
often a discrepancy between taxa targeted by studies wanting to use phylogenetic
information, and taxa for which phylogenies are available. Typically, the latter
result from focused studies of taxonomic groups, while the former encompass
species found in a given geographical location or ecosystem. We believe that by
providing an easy-to-use interface to obtain phylogenies for an arbitrary set of
taxa directly in R, \texttt{rotl} will be useful in a wide variety of contexts.

The accuracy and usefulness of the data provided by OTL relies on the community
to make generated phylogenies (and their metadata) digitally available as tree
files (i.e., Newick, NEXUS or NeXML). We strongly encourage researchers to
submit their published phylogenies to OTL using the curator interface
(\url{https://tree.opentreeoflife.org/curator}). By facilitating the discovery and
re-use of published trees and of the synthetic Open Tree, we hope \texttt{rotl} will
contribute to the wider adoption of best practices to make phylogenetic
information available and re-usable.

\section{Availability}
\label{sec:availability}

\texttt{rotl} is free, open source, and released under a Simplified BSD
license. Stable versions are available from the CRAN repository
(\url{https://cran.r-project.org/package=rotl}), and development versions are
available from GitHub (\url{https://github.com/ropensci/rotl}). This manuscript
was built using version 0.5.0.901-rotl-ms.2
(\url{https://github.com/ropensci/rotl/tree/v0.5.0.901-rotl-ms.2}). The package
is under active development, and authors welcome bug reports or feature requests
via the GitHub repository.  The source for this manuscript is available on
GitHub (\url{https://github.com/fmichonneau/rotl-ms}).

Python (\url{https://github.com/OpenTreeOfLife/pyopentree}) and Ruby
(\url{https://github.com/SpeciesFileGroup/bark}) libraries to interact
with the OTL APIs are also available.

\section{Acknowledgments}
\label{sec:acknowledgments}

We would like to thank the organizers of the OpenTree of Life APIs hackathon
that was held at the University of Michigan, Ann Arbor, 15-19 September, 2014,
where the development of \texttt{rotl} was started. We would also like to thank
Scott Chamberlain (rOpenSci) for providing a thorough code review, Scott
Chamberlain and Ross Mounce for commenting on the pre-print version of this
manuscript, and Shinichi Nakagawa and Alistair Senior for their help in
developing the package's meta-analysis vignette. DJW was supported by NIH Grant
R01-GM101352. FM was supported by iDigBio, and therefore this material is based
upon work supported by the National Science Foundation’s Advancing Digitization
of Biodiversity Collections Program (Cooperative Agreement EF-1115210).

\bibliographystyle{sysbio}

\bibliography{rotl-manuscript}

\end{document}